### 2008-11-16 papers/web_of_ideas/rebuttal.txt
Work on rebuttal

We would like to thank the reviewers for their comments. We believe that we should have little difficulty making the suggested changes.

The reviewers focussed on two related concerns: the relation of Think Link to prior work on argumentation in hypermedia systems, and the simple nature of our claim graph.

-- simple structure of the claim graph --

The AC is correct that the simple "supports/opposes" model that we use does not thoroughly model the semantic relation between claims; indeed it is not intended to. There is a trade-off between making a system rich enough to model complex relations between claims, and making it simple enough that it does not overwhelm users.

The intention of our claim graph is not to provide a rigorous graph of a logical argument, but to provide a user with a list of things that they might want to look at in order to assemble such an argument in their head. For example, if we look at Toulmin's standard example of asserting British Citizenship, Toulmin would say:

  claim: Harry is a British Citizen
  grounds: Harry was born in Bermuda 
  warrant: A man born in Bermuda will be a British Citizen

but Think Link would consider both the "grounds" and "warrant" to be "supporting" claims and leave it to the user to understand how to form a logical argument from these claims.

It is also useful to compare to ClaimMaker, which expands Think Link's "supports" into proves, isEvidenceFor, agreesWith, and isConsistentWith.

Our claim graph is also a simplification of that of IBIS. We merge "arguments" and "positions" into "claims", and merge "issues" into our "topics". Thus, while it is possible to map any IBIS graph into a Think Link graph, some type information will be lost. Like SynView, we extend IBIS by allowing users to rate the links they believe to be important.

We do not claim that our graph structure is optimal, or even necessarily better than the standard IBIS structure. Indeed we have informally experimented with a number of other graph structures, including having a separate "question" type (like IBIS), and introducing a "related" link type between claims.

The focus of the paper is the idea of allowing users to overlay such a structure over the existing web. It is quite likely that the details of the graph structure will evolve as we observe how Think Link is used.

TODO: talk about "refines" and "because" for claims.

-- relation to past work in hypermedia --

We agree with the AC that the paper should include a discussion of ZENO; indeed an earlier draft of the paper included a paragraph on ZENO which was accidentally cut.

As the reviewers note, that has been a lot of past work on using hypermedia systems to model arguments, and our paper would benefit from a more through commentary on its relation to this work.

Many of the early hypermedia systems (e.g. NoteCards, Xanadu, Intermedia, gIBIS, TextNet, and Augment) included typed links that could be used to model arguments. TextNet in particular provided a rich vocabulary of link types for such things as "refutes" and "supports".

Several hypermedia systems were explicitly designed to allow users to create structured arguments. For example gIBIS presented a graphical view of arguments created in the IBIS model,  *** TODO: read more about how these differ ***

Unlike the web, many of these hypermedia systems assumed that a linkable object would be a small snippet of text (around 100 words for NoteCards) rather than a large scrollable page. Indeed one can see Think Link's snippets as being an attempt to move the web back to this model.

Many early hypermedia systems also allowed a user to visually browse the graph of connected nodes (e.g. NoteCards, Intermedia, PlaneText, gIBIS, Belvedere, and Aquanet). Again, this is a concept that Think Link reintroduces to the web. When one clicks on a Think Link snippet one is shown a claim graph that bears some similarity to the graph browsers from these early hypertext systems.

The main differences between Think Link and this prior work result from it's application to the existing web, and the design implications that this has.

Most of the early hypertext argumentation systems.

One can see Think Link's snippets as being an attempt to move back to this model.

Other notable examples include EUCLID, TRELLIS, JANUS, AQUANET, BELVEDER, InterNote, and gIBIS.

TODO: in-depth description of this work...