### 2008-11-18 papers/web_of_ideas/rebuttal.txt
Final submitted version of the rebuttal

We would like to thank the reviewers for their comments. We believe that we should have little difficulty making the requested changes.

The reviewers focused on two related concerns: the simple nature of our claim graph, and the relationship of Think Link to prior work on argumentation in hypermedia systems.

-- simple claim graph --

There is a trade-off between making the claim graph rich enough to be useful, while making it simple enough that users are able to easy construct it. As the AC observes, our "supports/opposes" model ignores some of the details of argument structure. For example, if we look at Toulmin's standard example of asserting British citizenship, Toulmin would say:

  claim: Harry is a British Citizen
  grounds: Harry was born in Bermuda 
  warrant: A man born in Bermuda will be a British Citizen

Think Link considers the grounds and warrant to both be supporting claims. Toulmin puts more information in his graph, at the cost of requiring users to understand how to put arguments into this form. Think Link makes it easier to create the graph, but the graph contains less information and a user reading the graph needs to infer the logical connections between claims themselves.

Our intention is that Think Link allow users to identify things that are relevant to an argument for or against a claim, without requiring users to structure arguments an a rigorous logical way.

Our claim graph is a simplification of that of IBIS. IBIS distinguishes between "issues" (e.g. "what should we buy?"), "positions" (e.g. "we should buy a porsche") and arguments (e.g. "a porsche is fast"). We model both positions and arguments as claims. An IBIS issue can be either open (e.g. "what should we buy?"), or closed (e.g. "we should buy a porsche?"). We model an open issue as a topic, and a closed issue as a claim. Similarly all IBIS link types can be meaningfully mapped onto Think Link link types. Thus, while one can map any IBIS graph into a Think Link graph, some type information will be lost. Like SynView, we extend IBIS by allowing users to rate the links they believe to be important. We also extend IBIS by linking to evidence in the form of snippets. Note that the "because" weakness raised in our paper also applies to IBIS.

The focus of the paper is the idea of allowing users to overlay an argument structure over the existing web. We do not claim that our graph structure is optimal, or even necessarily better than IBIS. We have informally experimented with a number of other graph structures and it is likely that our graph structure will evolve as we learn from use in the field.

-- relation to past work in hypermedia --

We agree with the AC that the paper should include a discussion of ZENO; indeed an earlier draft of the paper included a paragraph on ZENO and IBIS which was mistakenly cut during a late compression pass.

As the reviewers note, there has been a lot of work on using hypermedia systems to model arguments, and we agree that our paper would benefit from a more through discussion of the relation of Think Link to this work - which we are happy to add.

Many of the early hypermedia systems (e.g. NoteCards, Xanadu, Intermedia, gIBIS, TextNet, and Augment) included typed links that could be used to model arguments. TextNet in particular provided a rich vocabulary of link types for such things as "refutes" and "supports". Several hypermedia systems were explicitly designed for argumentation (e.g. gIBIS, Belvedere, Aquanet, and TRELLIS).

Think Link's browsable claim graph and small snippets of text both hark back to earlier hypermedia systems. In many early hypermedia systems (e.g NoteCards) a linkable object is a small piece of text, rather than a large scrollable page. Similarly, many early hypermedia systems allow a user to visually browse the graph of connected nodes (e.g. NoteCards, Intermedia, PlaneText, gIBIS, Belvedere).

The key contribution that Think Link has over this past work is that we are layering a collaborative argument tool on top of the existing web, rather than creating a clean-slate argumentation tool. Think Link layers a new hypermedia system on top of an existing one. At the bottom, we have the existing web, in which scrollable pages are linked together according to the desires of the author. Above that we have a second hypermedia system in which snippets from the web pages are linked into a graph with claims and topics. Think Link thus provides users with a new, argument-focussed way to browse the web.

We would like to thank the AC and reviewers again for their help.