### 2009-10-10 papers/www_disputefinder/www_disputefinder.tex

\documentclass{www2010-submission}
\usepackage{graphics}
\usepackage{color}

\newcommand{\idea}[1]{\\\noindent{\color{blue} #1}\\}

\begin{document}
\title{Highlighting Disputed Claims on The Web}

\numberofauthors{5}

\author{
\alignauthor Rob Ennals\\
       \affaddr{Intel Labs Berkeley}\\
       \affaddr{2150 Shattuck Ave}\\
       \affaddr{Berkeley, CA, USA}\\
       \email{robert.ennals@intel.com}
\alignauthor Beth Trushkowsky\\
       \affaddr{Computer Science Division}\\
       \affaddr{University of California at Berkeley}\\
       \affaddr{Berkeley, CA, USA}\\
       \email{trush@berkeley.edu}
\alignauthor John Mark Agosta\\
       \affaddr{Intel Labs Santa Clara}\\
       \affaddr{2200 Mission College Blvd}\\
       \affaddr{Santa Clara, CA, USA}\\
       \email{john.m.agosta@intel.com}
}

\maketitle

\begin{abstract}
\idea{Dispute Finder highlights Disputed claims on the web}
\idea{Disputed means that a site you might trust gives a different point of view}
\idea{Use NLP techniques to highlight many pages that make the same claim}
\idea{Mine claims from existing sites, and from user submissions}
\idea{User testing discovers interesting issues, which directed the design}
\end{abstract}

\section{Introduction}

\idea{Web allows anyone to publish whatever they want. Sometimes this is good, but sometimes bad.}
\idea{Traditional newspapers are dying, or laying off fact checking staff. Can we provide an alterative.}

\idea{Two personas: Activist and Skeptical Reader}

\idea{Dispute Finder is available to the public.}
\idea{As yet, around 12k people have installed it and tried it out, but of those, only around 2k remain active daily users and around 1k actively contribute. Number of users rises and falls with media attention. }

\section{Related Work}

\subsection{Checking Accuracy and Bias on the Web}

\idea{Snops, Fact Check, Politifact, etc}

\idea{NewsCube, MediaCloud, etc}

\subsection{Web Annotation and Clipping Systems}

\idea{Hunter Gatherer}

\idea{VideoLyzer}

\end{document}