### 2010-01-26 papers/wicow_claimfinder/relatedwork.tex
Split related work out into a separate file to make it easier to edit independently

\subsection{What is credible?}

\subsection{The distributional hypothesis}

\subsection{Contradiction Detection}

\subsection{Sentiment Analysis}

\subsection{Crowdsourcing disputed claims}

\x{Dispute Finder. Wikipedia. WikiTrust}

\subsection{Curated Databases of Disputed Claims}

\x{Dispute Finder.}