### 2010-01-28 papers/wicow_claimfinder/relatedwork.tex
No changes. Commiting to allow merge.

\todo{For the moment we put related work in this section. Most {\it background} related work should probably be included directly in the paper body rather than in the front section.}

\subsection{What is credible?}

\subsection{The distributional hypothesis}

\subsection{Contradiction Detection}

\subsection{Sentiment Analysis}

\subsection{Crowdsourcing disputed claims}

\x{Dispute Finder. Wikipedia. WikiTrust}

\subsection{Curated Databases of Disputed Claims}

\x{Dispute Finder.}